Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response

### 2024-12-11 paper/paper.qmd
Add datasheet section to paper

---
title: "Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response"
subtitle: "Identifying Key Factors for Fire Severity and Evaluating the Effectiveness of Public Fire Response Systems"
author: 
  - Maggie Zhang
thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
date: "December 3, 2024"
date-format: long
abstract: "This study investigates fire severity levels in Toronto to improve resource allocation for Toronto Fire Services (TFS). The analysis identifies that higher-severity incidents are strongly correlated with incidents originating from mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires. The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment, are more likely to be severe. The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are successful in low and medium-severity fires, while sprinkler systems are often only partially operational. Results can guide TFS to enhance response times, prioritize resources, and optimize fire response operations, ultimately improving community safety in Toronto."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows. [@sec-data] describes the data used in the analysis, including the sources and key variables and their visualizations.[@sec-model] outlines the models applied to estimate fire severity and identify key factors. [@sec-results] presents the results of the analysis. [@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements. Finally, [@sec-appendix] includes supplementary information and additional data details.

@@

This structured approach allows us to analyze and understand various aspects of fire incidents in Toronto. All variables in the dataset have the potential to provide valuable insights and perspectives for various analyses, contributing to fire prevention and safety improvements. However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study. A more detailed explanation will be provided in the @sec-appendix.

@@

**Understanding Variables of Interest**:\
To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics. The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling. Can check @sec-appendix for visualization/summary of the distribution of all variables. Below are a few important variables considered:

@@

2.  **Civilian Casualties**:\
    This is a numerical variable that provides insight into the severity of the incidents. Basic descriptive statistics (e.g., mean, median, and range) were calculated using the `summary()` function. Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see @tbl-civilian-casualties-summary). We can see that most of the time, there are no casualties. This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendix), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.

@@

# Variable Selection and Visualization

@@

# Idealized Research Methodology

>>>>

---
title: "Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response"
subtitle: "Identifying Key Factors for Fire Severity and Evaluating the Effectiveness of Public Fire Response Systems"
author: 
  - Maggie Zhang
thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
date: today
date-format: long
abstract: "This study investigates fire severity levels in Toronto to improve resource allocation for Toronto Fire Services (TFS). The analysis identifies that higher-severity incidents are strongly correlated with incidents originating from mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires. The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment, are more likely to be severe. The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are successful in low and medium-severity fires, while sprinkler systems are often only partially operational. Results can guide TFS to enhance response times, prioritize resources, and optimize fire response operations, ultimately improving community safety in Toronto."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows. [@sec-data] describes the data used in the analysis, including the sources and key variables and their visualizations.[@sec-model] outlines the models applied to estimate fire severity and identify key factors. [@sec-results] presents the results of the analysis. [@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements. Finally, [@sec-appendixa] includes details about variable selection and cleaning, with the visualization of all variables in the analysis dataset; [@sec-appendixb] is a supplementary datasheet that explains the building of the analysis dataset; [@sec-appendixc] outlines the idealized methodology for conducting this study, given a budget of $100K.

@@

This structured approach allows us to analyze and understand various aspects of fire incidents in Toronto. All variables in the dataset have the potential to provide valuable insights and perspectives for various analyses, contributing to fire prevention and safety improvements. However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study. A more detailed explanation will be provided in the @sec-appendixa.

@@

**Understanding Variables of Interest**:\
To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics. The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling. Can check @sec-appendixa for visualization/summary of the distribution of all variables. Below are a few important variables considered:

@@

2.  **Civilian Casualties**:\
    This is a numerical variable that provides insight into the severity of the incidents. Basic descriptive statistics (e.g., mean, median, and range) were calculated using the `summary()` function. Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see @tbl-civilian-casualties-summary). We can see that most of the time, there are no casualties. This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.

@@

# Variable Selection and Visualization {#sec-appendixa}

@@

# Supplemental Datasheet {#sec-appendixb}

A datasheet for the dataset used in this analysis is included with this paper. Please
refer to this link to the [Datasheet](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident/blob/main/paper/datasheet.pdf).

# Idealized Research Methodology {#sec-appendixc}
